---
title: "R Notebook"
output: html_notebook
---

```{r}
# load libraries 
library(dplyr)
options(dplyr.summarise.inform = FALSE)
library(tidyr)
```

```{r}
# load samples
load('data/waves.Rdata')
# load census tables 
load('data/poptable.Rdata')
```

```{r}
# import functions
source('helper_rev.R')
```


```{r}
# define fixed objects
coalitions <- c('CSX', 'CDX', 'Az', 'M5S', 'altri')
# define available adjustment columns
COLS <-  c('gen', 'age', 'emp', 'edu', 'zone', 'pastvote')
# define census tables to iterate over
census_list <- c('edu', 'emp', 'pastvote', 'approx')

# build iteration grid
iter_grid <- expand.grid(
  df_name = names(main),
  census_name = census_list,
  stringsAsFactors = FALSE
)

# run one sample-census combination
run_one <- function(df_name, census_name, iter_id = NA, n_iter = NA) {
  # print progress
  print(paste('processing', iter_id, 'of', n_iter, df_name, census_name))
  # select census 
  ref_table <- poptable[[census_name]]
  # select data 
  to_adj <- main[[df_name]]
  # check or pastvote 
  census_cols <- colnames(ref_table)[colnames(ref_table) %in% COLS]
  # check available sample columns
  data_cols <- colnames(to_adj)[colnames(to_adj) %in% COLS]
  
  # skip sample-census combinations that cannot be adjusted
  if (!all(census_cols %in% data_cols)) {
    # print skipped combination
    print(paste('skipping', df_name, census_name, '- missing columns in sample'))
    # return empty result
    return(NULL)
  }
  
  # unweighted
  unweighted_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'unadjusted',
    uniweighted(to_adj)
  )

  # post-stratification or calibration (the same here)
  poststrat_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'post-stratification',
    poststrat(to_adj, ref_table)
  )

  # multinomial linear model-assisted post-stratification
  multinomial_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'MN post-stratification',
    multinomial(
      to_adj,
      ref_table,
      B = 5 # 1000
    )
  )

  # generate approximate reference sample
  reference_survey <- generate_reference(census = ref_table)
  # logit IPW
  ipw_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'Logit-IPW',
    ps_ipw(to_adj, reference_survey, trim = 5, B = 5)
  )

  # # generate approximate reference sample
  # reference_survey <- generate_reference(census = ref_table)
  # run IPW with a gradient-boosted selection model
  xgboost_out <- nonprob_ipw_xgb(
    # non-probability survey
    data = to_adj,
    # generated reference sample
    reference_survey = reference_survey,
    # bootstrap SEs
    var_method = "bootstrap",
    # bootstrap repetitions
    num_boot = 10 # 1000
  )
  # XGBoost IPW
  xgboost_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'XGboost-IPW',
    xgboost_out
  )

  # DRP
  drp_out <- data.frame(
    sample = df_name,
    census = census_name,
    method = 'DRP',
    nonprob_drp(to_adj, ref_table)
  )

  # # force rstan to avoid parallel chains
  # options(mc.cores = 1)
  # mrp_out <- mrp(
  #   data = to_adj,
  #   census = ref_table,
  #   draws = 50,
  #   chains = 2,
  #   iter = 200,
  #   cores = 1
  # )
  # # MPR
  # mrp_out <- data.frame(sample = df_name,
  #                   census = census_name,
  #                   method = 'MRP',
  #                   mrp_out
  # )
  
  # # NN
  # nn_out <- nn_poststrat(
  #   data = to_adj,
  #   census = ref_table,
  #   B = 2, # 1000
  #   n_epochs = 50, 
  #   n_neurons = 2, #8,
  #   n_hidden = 2, 
  #   batch_size = 32,
  #   learning_rate = 1e-3,
  #   early_stopping = TRUE
  # )
  # nn_out <- data.frame(sample = df_name,
  #                   census = census_name,
  #                   method = 'NN',
  #                   nn_out
  # )
  
  # append all results for this sample-census combination
  rbind(
    unweighted_out,
    poststrat_out,
    multinomial_out,
    ipw_out,
    xgboost_out,
    drp_out
    # mrp_out,
    # nn_out
  )
}

# run all iterations using lapply
out_list <- lapply(seq_len(nrow(iter_grid)), function(i) {
  # run one sample-census combination
  run_one(
    df_name = iter_grid$df_name[i],
    census_name = iter_grid$census_name[i],
    iter_id = i,
    n_iter = nrow(iter_grid)
  )
})

# append the results into a common dataframe
out <- do.call(rbind, out_list)

# clean row names
rownames(out) <- NULL
```


```{r}
# MRP 
load('results/MRP.Rdata')
# XGBoost
load('results/XGBoost.Rdata')
# others
load('results/others.Rdata')
# NN 
load('results/NN_10.Rdata')
```


```{r}
# check that outcomes sum to 1 
out %>% group_by(sample, census, method) %>% summarise(test = round(sum(estimate), 4) != 1) %>% filter(test)
# out %>% filter(method == 'Logit-IPW')
```



```{r}
# unweighted
uniweighted(main$wave1)

# post-stratification or calibration (the same here)
poststrat(main$wave1, poptable$pastvote) 
# calib(main$wave1, poptable$pastvote, calfun = "linear") 
# calib(main$wave1, poptable$pastvote, calfun = "logit") 

# multinomial
multinomial(main$wave1, poptable$pastvote, 
            B = 30 # 1000
            )

# generate approximate reference sample
reference_survey <- generate_reference(census = poptable$census)
# run IPW
nonprob_ipw(
  # non-probability survey
  data = main$wave1,
  # generated reference sample
  reference_survey = reference_survey, 
  # bootstrap SEs
  var_method = "bootstrap",
  # bootstrap repetitions
  num_boot = 30 # 1000
  # var_method = "analytic" # <- returns the same value of se for all parties
)

# run DRP
nonprob_drp(data = main$wave2, census = poptable$edu)

# force rstan to avoid parallel chains
options(mc.cores = 1)
mrp(
  data = main$wave1,
  census = poptable$census,
  draws = 50,
  chains = 2,
  iter = 200,
  cores = 1
)

# NN 
nn_poststrat(
  data = main$wave1,
  census = poptable$census,
  B = 10,
  n_epochs = 50,
  n_neurons = 8,
  n_hidden = 2,
  batch_size = 32,
  learning_rate = 1e-3,
  early_stopping = TRUE
)

# generate approximate reference sample
reference_survey <- generate_reference(census = poptable$edu)
# run IPW with a gradient-boosted selection model
nonprob_ipw_xgb(
  # non-probability survey
  data = main$wave4,
  # generated reference sample
  reference_survey = reference_survey,
  # bootstrap SEs
  var_method = "bootstrap",
  # bootstrap repetitions
  num_boot = 2 # 1000
)
```

# check weights in IPW estimation

```{r}
# refit once and keep the underlying nonprob object
diag <- nonprob_ipw(data = main$wave4, reference_survey = generate_reference(census = poptable$edu), return_fit = TRUE)
# the fitted model
fit <- diag$fit
# find the propensity / weight slots (names vary by version)
names(fit)
# print built-in diagnostics
summary(fit)

# inverse-probability weights
w <- 1 / fit$ps_scores
# the handful of giant weights is the smoking gun
sort(w, decreasing = TRUE)[1:10]

hist(fit$ps_scores)
```


# ------------------------------------------------------------
# ---- check different selection models to justify choice ----
# ------------------------------------------------------------
```{r}
# choose predictors
preds <- c("age", "gen", "edu", "emp", "zone", "pastvote")

# quick weighted AUC
w_auc <- function(y, p, w) {
  # keep valid rows
  ok <- complete.cases(y, p, w)
  y <- y[ok]; p <- p[ok]; w <- w[ok]
  
  # order predictions
  o <- order(p)
  y <- y[o]; p <- p[o]; w <- w[o]
  
  # compute pairwise weighted AUC
  neg_below <- 0; num <- 0
  for (i in split(seq_along(p), p)) {
    pos_w <- sum(w[i][y[i] == 1])
    neg_w <- sum(w[i][y[i] == 0])
    num <- num + pos_w * (neg_below + 0.5 * neg_w)
    neg_below <- neg_below + neg_w
  }
  
  # normalise by possible positive-negative pairs
  den <- sum(w[y == 1]) * sum(w[y == 0])
  ifelse(den > 0, num / den, NA_real_)
}

# prepare sample cases
samp <- main$wave1 %>%
  # keep complete predictors
  filter(complete.cases(across(all_of(preds)))) %>%
  # harmonise classes
  mutate(across(all_of(preds), as.character),
         S = 1,
         w = 1) %>%
  # keep only model columns
  select(all_of(preds), S, w)

# prepare population/reference cases
pop <- poptable$census %>%
  # keep complete predictors and positive cells
  filter(complete.cases(across(all_of(c(preds, "count")))),
         count > 0) %>%
  # harmonise classes
  mutate(across(all_of(preds), as.character)) %>%
  # aggregate possible duplicate cells
  group_by(across(all_of(preds))) %>%
  summarise(count = sum(count), .groups = "drop") %>%
  # balance total reference weight to sample size
  mutate(S = 0,
         w = count / sum(count) * nrow(samp)) %>%
  # keep only model columns
  select(all_of(preds), S, w)

# combine sample and reference data
dat <- bind_rows(samp, pop) %>%
  # factors for categorical predictors
  mutate(across(all_of(preds), as.factor),
         age_n = as.numeric(age))

# create stratified CV folds
set.seed(123)
dat$fold <- NA_integer_
for (s in 0:1) {
  idx <- which(dat$S == s)
  dat$fold[idx] <- sample(rep(1:5, length.out = length(idx)))
}

# benchmark additive formula
main_rhs <- paste(preds, collapse = " + ")

# candidate selection models
forms <- c(
  main = paste("S ~", main_rhs),
  age_gen = paste("S ~", main_rhs, "+ age:gen"),
  age_edu = paste("S ~", main_rhs, "+ age:edu"),
  gen_edu = paste("S ~", main_rhs, "+ gen:edu"),
  zone_pastvote = paste("S ~", main_rhs, "+ zone:pastvote"),
  age_quadratic = "S ~ age_n + I(age_n^2) + gen + edu + emp + zone + pastvote"
)

# score one formula
score_model <- function(f) {
  # fit full model for parameter count
  fit_full <- suppressWarnings(glm(as.formula(f), dat, weights = w, family = binomial()))
  
  # create CV predictions
  p <- rep(NA_real_, nrow(dat))
  for (k in 1:5) {
    fit_k <- suppressWarnings(glm(as.formula(f), dat[dat$fold != k, ], weights = w, family = binomial()))
    p[dat$fold == k] <- predict(fit_k, dat[dat$fold == k, ], type = "response")
  }
  
  # return compact result
  data.frame(
    n_par = sum(!is.na(coef(fit_full))),
    cv_auc = w_auc(dat$S, p, dat$w)
  )
}

# run comparison
ipw_auc_check <- bind_rows(lapply(names(forms), function(m) {
  cbind(model = m, score_model(forms[m]))
})) %>%
  # compare against additive benchmark
  mutate(
    added_par = n_par - n_par[model == "main"],
    auc_gain = cv_auc - cv_auc[model == "main"]
  ) %>%
  # best improvements first
  arrange(desc(auc_gain))

# final comparison
ipw_auc_check
```


# table of correlation strenght, n of cells, distribution difference 

```{r}
# libraries (loaded, never namespace-qualified)
library(dplyr)
library(tidyr)
library(nnet)

# load the survey microdata (df: covariates + outcome nowvote)
load('data/waves.Rdata')
# load the population tables (poptable: waves + census margins)
load('data/poptable.Rdata')
```


```{r}
# census tables to evaluate
census_list <- c('edu', 'emp', 'pastvote', 'approx')
# detect the wave samples actually present (note: there is no wave4)
wave_names <- grep('^wave', names(poptable), value = TRUE)

# the common covariates of a census table are all its columns except the frequency
covs_of <- function(ct) setdiff(names(ct), 'count')

# ---- metric 1: out-of-sample proxy strength (k-fold CV correlation) ----
strength_fun <- function(data, covs, outcome = 'nowvote', k = 5, seed = 1) {
  # reproducible folds
  set.seed(seed)
  # keep only the modelling columns and complete rows
  d <- data[c(covs, outcome)] %>% filter(complete.cases(.))
  # outcome as a factor plus its full level set
  y <- factor(d[[outcome]]); lev <- levels(y)
  # one-hot encoding of the true labels (target of the correlation)
  onehot <- model.matrix(~ y - 1)
  # assign each row to one of k folds
  fold <- sample(rep(1:k, length.out = nrow(d)))
  # container for out-of-sample predicted class probabilities
  pred <- matrix(0, nrow(d), length(lev), dimnames = list(NULL, lev))
  # multinomial formula from the census-common covariates
  form <- as.formula(paste(outcome, '~', paste(covs, collapse = ' + ')))
  # loop over folds, training on the rest and predicting the held-out part
  for (f in 1:k) {
    # split train / test for this fold
    tr <- d[fold != f, ]; te <- d[fold == f, ]
    # fit the multinomial proxy model (quiet)
    m <- multinom(form, data = tr, trace = FALSE, maxit = 200)
    # predict held-out class probabilities
    p <- predict(m, newdata = te, type = 'probs')
    # guard the 2-class case where predict() returns a vector
    if (is.null(dim(p))) p <- cbind(`1` = 1 - p, `2` = p)
    p <- as.matrix(p)
    # add any class missing from this fold as a zero column
    miss <- setdiff(lev, colnames(p))
    if (length(miss)) p <- cbind(p, matrix(0, nrow(p), length(miss), dimnames = list(NULL, miss)))
    # store the predictions aligned to the full level order
    pred[fold == f, lev] <- p[, lev]
  }
  # single correlation score between predicted probs and true one-hot labels
  cor(as.vector(pred), as.vector(onehot))
}

# ---- metric 2: average cell discrepancy (wave vs census) ----
discrepancy_fun <- function(wave_tbl, census_tbl) {
  # covariates shared between this census and the survey cells
  covs <- covs_of(census_tbl)
  # collapse the wave to the census covariates -> observed cell frequencies
  obs <- wave_tbl %>%
    ungroup() %>%
    group_by(across(all_of(covs))) %>%
    summarise(f_data = sum(count), .groups = 'drop')
  # rename the census frequency for a clean join
  cen <- census_tbl %>% rename(f_cens = count)
  # align both distributions on the common cells and average the gap
  obs %>%
    full_join(cen, by = covs) %>%
    # cells missing on either side contribute zero frequency
    mutate(f_data = coalesce(f_data, 0), f_cens = coalesce(f_cens, 0)) %>%
    # mean absolute discrepancy across cells
    summarise(discrepancy = mean(abs(f_data - f_cens))) %>%
    pull(discrepancy)
}

# ---- census-level quantities (strength + n_cells depend on the census only) ----
census_tbl <- tibble(census = census_list) %>%
  # one row per census, computed once
  rowwise() %>%
  mutate(
    # CV proxy strength of this census' covariates for nowvote
    strength = strength_fun(df, covs_of(poptable[[census]])),
    # number of post-stratification cells defined by this census
    n_cells = nrow(distinct(poptable[[census]][covs_of(poptable[[census]])]))
  ) %>%
  ungroup()

# ---- assemble the full (sample x census) grid ----
results <- expand_grid(sample = wave_names, census = census_list) %>%
  # discrepancy is the only quantity that varies by wave
  rowwise() %>%
  mutate(discrepancy = discrepancy_fun(poptable[[sample]], poptable[[census]])) %>%
  ungroup() %>%
  # attach the census-level strength and n_cells
  left_join(census_tbl, by = 'census') %>%
  # final column order
  select(sample, census, strength, discrepancy, n_cells)

# inspect
results
# save
save(results, file = 'saves/proxy_metrics.Rdata')
```





# ------------------------------------------------------------
# ------------------------ older code ------------------------
# ------------------------------------------------------------



# plotting for Survey Research Methods paper 

```{r}
# convert wave into dataset and order methods - necessary to make plot
method_conv <- c("raking" = 'Raking', "PS-IPW" = "PS-IPW", "post-strat" = 'MN', "DRP" = 'DRP', "MRP" = 'MRP', "NN" = 'NN')
to_plot$method <- method_conv[to_plot$method]
to_plot$method <- factor(to_plot$method, levels = c('Raking', 'PS-IPW', 'MN', 'NN', 'MRP', 'DRP'))
wave_conv <- c("wave1" = 'dataset 1', "wave2" = 'dataset 2', "wave3" = 'dataset 3', 
               "wave4" = 'dataset 4', "wave5" = 'dataset 5')
to_plot$dataset <- wave_conv[to_plot$wave]
to_plot$wave <- NULL
to_plot$table <- factor(to_plot$table, levels = c('edu', 'emp', 'pastvote', 'approx', 'marginals'))
# save(to_plot, file = 'results/to_plot_named.Rdata')
```

```{r}
to_table <- to_plot %>% group_by(method) %>% summarise(tot_se = sum(stats.se))
to_table
```


```{r}
library(ggplot2)
se_plot <- to_plot %>% group_by(method, table, dataset) %>% summarise(tot_var = sum(stats.se^2),
                                          tot_bias = sum(abs(stats.A - stats.U)))
ggplot(se_plot, aes(x = method, y = tot_bias, shape = table, color = dataset, linetype = table, group = interaction(table, dataset))) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  theme_minimal(base_size = 14) + 
  labs(x = '', y = 'cross-party bias', subtitle =
         paste0('strat_tables_1000, unadj_mse.Rdata, marginals_raking_1000, pre.Rmd, ', Sys.Date())) + 
  theme(plot.subtitle = element_text(size = 8)) 

# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/total_bias.pdf')
```

```{r}
se_plot <- to_plot %>% group_by(method, table, dataset) %>% summarise(tot_se = sum(stats.se),
                                          tot_bias = sum(abs(stats.A - stats.U)))
ggplot(se_plot, aes(x = method, y = tot_se, shape = table, color = dataset, linetype = table, group = interaction(table, dataset))) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  theme_minimal(base_size = 14) + 
  labs(x = '', y = 'cross-party SE', subtitle =
         paste0('strat_tables_1000, unadj_mse.Rdata, marginals_raking_1000, pre.Rmd, ', Sys.Date())) + 
  theme(plot.subtitle = element_text(size = 8)) 

# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/total_se.pdf')
```



```{r}
library(ggplot2)
to_gplot <- to_plot 
to_gplot$comb <- apply(to_gplot[, c('dataset', 'table')], 1, function(x) paste(x, collapse = ' '))
ggplot(to_gplot, aes(x = method, y = rmse_B, shape = table, color = dataset, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  scale_x_discrete(expand = c(0.033, 0.033)) +
  theme_minimal(base_size = 12) + 
  labs(x = '', y = 'RMSE B', subtitle =
         paste0('strat_tables_1000, unadj_mse.Rdata, marginals_raking_1000, pre.Rmd, ', Sys.Date())) +
  theme(plot.subtitle = element_text(size = 8)) 
# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/rmse_B.pdf')
# save in THESIS
# ggsave('/Users/albertoarletti/Documents/Overleaf/Template PhD Thesis/Figures/mse_B.pdf')
```


```{r}
ggplot(to_gplot, aes(x = method, y = rmse_wB, shape = table, color = dataset, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  scale_x_discrete(expand = c(0.033, 0.033)) +
  theme_minimal(base_size = 12) + 
  labs(x = '', y = 'RMSE wB', subtitle =
         paste0('strat_tables_1000, unadj_mse.Rdata, marginals_raking_1000, pre.Rmd, ', Sys.Date())) +
  theme(plot.subtitle = element_text(size = 8)) 
# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/rmse_wB.pdf')
# save in THESIS
# ggsave('/Users/albertoarletti/Documents/Overleaf/Template PhD Thesis/Figures/mse_wB.pdf')
```




```{r}
library(ggplot2)
to_gplot <- to_plot 
to_gplot$comb <- apply(to_gplot[, c('dataset', 'table')], 1, function(x) paste(x, collapse = ' '))
ggplot(to_gplot, aes(x = method, y = B, shape = table, color = dataset, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  scale_x_discrete(expand = c(0.033, 0.033)) +
  theme_minimal(base_size = 12) + 
  labs(x = '', y = 'Bias B', subtitle =
         paste0('results/strat_tables_tmp_approx, strat_tables_tmp, marginals_raking, pre.Rmd, ', Sys.Date())) + 
  theme(
    plot.subtitle = element_text(size = 8),
    # legend.position = "bottom"  # Move the legend to the bottom
  ) # +
  # guides(
  #   color = guide_legend(nrow = 2),    # Single row for the `dataset` legend
  #   shape = guide_legend(nrow = 2),    # Single row for the `table` legend
  # )
# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/B.pdf')
# save in THESIS
ggsave('/Users/albertoarletti/Documents/Overleaf/Template PhD Thesis/Figures/bias_B.pdf')

ggplot(to_gplot, aes(x = method, y = wB, shape = table, color = dataset, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  scale_x_discrete(expand = c(0.033, 0.033)) +
  theme_minimal(base_size = 12) + 
  labs(x = '', y = 'Bias wB', subtitle =
         paste0('results/strat_tables_tmp_approx, strat_tables_tmp, marginals_raking, pre.Rmd, ', Sys.Date())) + 
  theme(
    plot.subtitle = element_text(size = 8),
    # legend.position = "bottom"  # Move the legend to the bottom
  ) # +
  # guides(
  #   color = guide_legend(nrow = 2),    # Single row for the `dataset` legend
  #   shape = guide_legend(nrow = 2),    # Single row for the `table` legend
  # )
# ggsave('/Users/albertoarletti/Documents/Overleaf/SRM LaTeX Template/assets/wB.pdf')
# save in THESIS
ggsave('/Users/albertoarletti/Documents/Overleaf/Template PhD Thesis/Figures/bias_wB.pdf')
```

```{r}
ggplot(to_gplot, aes(x = method, y = B, shape = table, color = dataset, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2.5) +
  geom_line(linewidth = 0.5, alpha = 0.45) + 
  theme_minimal(base_size = 12) + 
  labs(x = '', y = 'B', subtitle =
         paste0('results/strat_tables_tmp_approx, strat_tables_tmp, marginals_raking, pre.Rmd, ', Sys.Date())) + 
  theme(
    plot.subtitle = element_text(size = 8),
    legend.position = "bottom"  # Move the legend to the bottom
  ) +
  guides(
    color = guide_legend(nrow = 2),    # Single row for the `dataset` legend
    shape = guide_legend(nrow = 2),    # Single row for the `table` legend
  )
```



# plotting and presentation 

```{r}
method_sel <- 'Raking'
wave_sel <- 'wave1'
table_sel <- 'edu'
to_gplot <- to_plot %>% filter(method == method_sel, wave == wave_sel, table == table_sel)
to_gplot$party <- factor(to_gplot$party, levels = coalitions)
ggplot(to_gplot, aes(x = party)) + 
  geom_point(aes(y = stats.A, shape = "Adjusted")) +  
  geom_point(aes(y = stats.U, shape = "Unadjusted")) +
  geom_errorbar(aes(ymin = stats.A - (2*stats.se), ymax = stats.A + (2*stats.se)), width = 0.2) + 
  geom_point(aes(y = poptable$GT, shape = "Ground Truth"), size = 2) +
  labs(y = '', 
       title = 'Estimated population values',
       subtitle = paste0('method: ', method_sel, '\n', 
                         'dataset: ', wave_sel, '\n', 'post-strat: ', table_sel)) +
  scale_shape_manual(values = c("Adjusted" = 16, "Unadjusted" = 2, "Ground Truth" = 3)) +
  theme(legend.title = element_blank(), legend.position = "right")  # Adjust legend position and title
```

```{r}
unique(to_plot$table)
```

# generate latex tables fro standard errors and confidence intervals 
```{r}
library(tidyverse)
table_chosen <- 'approx'
# Define the path of the file
  file_path <- "/Users/albertoarletti/Desktop/output.txt"
  # Open the file for writing
  file_conn <- file(file_path, "w")
for (party in coalitions){
  party_chosen <- party
  to_table <- to_plot %>% filter(table == table_chosen, party == party_chosen)
  stage1 <- to_table[c('method', 'wave', 'stats.se')] %>% pivot_wider(names_from = wave, values_from = stats.se)
  stage2 <- to_table[c('method', 'wave', 'stats.in_conf')] %>% pivot_wider(names_from = wave, values_from = stats.in_conf)
  tmp <- stage1$method
  stage2$method <- NULL
  stage1$method <- NULL
  # Combine data frames
  combined_df <- do.call(cbind, lapply(names(stage1), function(col) {cbind(stage1[[col]], stage2[[col]])}))
  colnames(combined_df) <- as.vector(t(outer(names(stage1), c("_se", "_conf"), paste0)))
  combined_df <- as.data.frame(combined_df)
  combined_df$method <- tmp
  combined_df <- combined_df[colnames(combined_df)[c(length(colnames(combined_df)), 1:(length(colnames(combined_df))-1))]]
  combined_df[, 2:length(colnames(combined_df))] <- round(combined_df[, 2:length(colnames(combined_df))] %>% mutate_all(as.numeric), 4)
  # combined_df[, 2:length(colnames(combined_df))]
  tmp <- as.data.frame(apply(combined_df, c(1, 2), function(x) if(x == 1){'$\\bullet$'}else{if (x == 0){''}else{x}}))
  tmp2 <- apply(tmp, 1, function(x) paste(x, collapse = ' & '))
  
  # Use a for loop to write each line
  part1_text <- "
  \\begin{table}[h!]
      \\centering
      \\resizebox{1\\textwidth}{!}{%
      \\begin{tabular}{lcccccccccc}
      \\hline
      \\textbf{Method} & \\multicolumn{2}{c}{\\textbf{Dataset 1}} & \\multicolumn{2}{c}{\\textbf{Dataset 2}} & \\multicolumn{2}{c}{\\textbf{Dataset 3}} & \\multicolumn{2}{c}{\\textbf{Dataset 4}} & \\multicolumn{2}{c}{\\textbf{Dataset 5}} \\\\
      \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9} \\cmidrule(lr){10-11}
       & \\textbf{se} & \\textbf{in conf.} & \\textbf{se} & \\textbf{in conf.} & \\textbf{se} & \\textbf{in conf.} & \\textbf{se} & \\textbf{in conf.} & \\textbf{se} & \\textbf{in conf.} \\\\
       \\hline
  "
  part2_text <- "
      \\end{tabular}
      }
      \\caption{Value of the estimated standard error for each method and dataset, for \\textbf{'sostituiscimi_party'} party only,  \\textbf{'sostituiscimi_table'} post-stratification table. The black bullets indicate whether the true population is included in the interval.}
      \\label{tab:se_inconf_sostituiscimi_party_sostituiscimi_table}
  \\end{table}
  "
  part2_text <- gsub("sostituiscimi_table", table_chosen, part2_text)
  part2_text <- gsub("sostituiscimi_party", party_chosen, part2_text)
  
  writeLines(part1_text, file_conn)
  for (row in tmp2){
    # Create a new line of text
    line <- paste(row, "\\\\ ")
    # Write the line to the file
    writeLines(line, file_conn)
  }
  writeLines(part2_text, file_conn)
  # Close the file connection
}
close(file_conn)
```



```{r}
load('results/pre_chapter2.Rdata')
```

```{r}
# to_save$out
to_plot[to_plot$method %in% c('NN', 'MRP'), ] <- to_save$out
```

```{r}
# main statistics 
to_plot %>% group_by(method) %>% summarise(b = mean(stats.b), 
                                           B = mean(B), 
                                           wB = mean(wB), 
                                           in_conf = mean(stats.in_conf), 
                                           se = mean(stats.se)) %>% 
  mutate(across(where(is.numeric), round, digits = 4))
```


```{r}
method_conv <- c("raking" = 'Raking', "PS-IPW" = "PS-IPW", "post-strat" = 'MN', "DRP" = 'DRP', "MRP" = 'MRP', "NN" = 'NN')
to_plot$method <- method_conv[to_plot$method]
to_plot$method <- factor(to_plot$method, levels = c('Raking', 'PS-IPW', 'MN', 'NN', 'MRP', 'DRP'))
```


```{r}
wave_conv <- c("wave1" = 'dataset 1', "wave2" = 'dataset 2', "wave3" = 'dataset 3', 
               "wave4" = 'dataset 4', "wave5" = 'dataset 5')
to_plot$wave <- wave_conv[to_plot$wave]
```

```{r}
library(ggplot2)
to_gplot <- to_plot 
tmp <- to_plot %>% group_by(method) %>% summarise('value' = mean(B))
ggplot(to_gplot, aes(x = method, y = B, shape = table, color = wave)) + 
  # ggrain::geom_rain(fill = 'grey') + 
  # theme_minimal() +
  # labs(y = 'b', subtitle = paste0('pre.Rmd, ', Sys.Date()))
  geom_point(alpha = 1, size = 2) +
  theme_minimal(base_size = 12) +
  # theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1)) +
  geom_point(data = tmp, aes(x = method, y = value, color = 'average'), size = 3, shape = 4, inherit.aes = FALSE) +
  # geom_line(data = tmp, aes(x = method, y = value, group = 1, color = 'average'), linetype = "dashed", inherit.aes = FALSE) +
  scale_color_manual(
    values = c(setNames(scales::hue_pal()(length(unique(to_gplot$wave))), unique(to_gplot$wave)), "average" = "black"),
    breaks = c(unique(to_gplot$wave), "average")
  ) +
  labs(color = "dataset", shape = 'post-strat table', x = '', y = 'B', subtitle =
         paste0('results/strat_tables_tmp_approx.Rdata, results/strat_tables_tmp.Rdata, pre.Rmd, ', Sys.Date())) + theme(plot.subtitle=element_text(size=8))
# ggsave('plots/B_alt.png')

```
# best method for each combination 
```{r}
tmp <- to_plot %>% group_by(wave, table, method) %>% summarise(B = mean(B)) 
tmp2 <- tmp %>% group_by(wave, table) %>% summarise(w = tmp$method[which(B == max(B))])
table(tmp2$w)
```

# count how many times each method had a B > 0 
```{r}
# total is (5 * 3) + 4 = 19 
# obtain frequency 
tmp <- to_plot %>% group_by(wave, table, method) %>% summarise(B = mean(B), wB = mean(wB)) 
tmp %>% group_by(method) %>% summarise('% B > 0' = sum(B > 0) / n(), '# B > 0' = sum(B > 0), mean(B), sd(B), '% wB > 0' = sum(wB > 0) / n(), '# wB > 0' = sum(wB > 0), mean(wB), sd(wB)) %>% mutate(across(where(is.numeric), function(x) round(x, 2)))
```


# plot with lines connecting same combination 
```{r}
library(ggplot2)
to_gplot <- to_plot # %>% filter(wave == 'wave2')
to_gplot$comb <- apply(to_gplot[, c('wave', 'table')], 1, function(x) paste(x, collapse = ' '))
ggplot(to_gplot, aes(x = method, y = B, shape = table, color = wave, group = comb, linetype = table)) + 
  geom_point(alpha = 1, size = 2) + 
  geom_line() + 
  theme_minimal(base_size = 12)
#   # theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1)) +
#   geom_point(data = tmp, aes(x = method, y = value, color = 'average'), size = 3, shape = 4, inherit.aes = FALSE) +
#   # geom_line(data = tmp, aes(x = method, y = value, group = 1, color = 'average'), linetype = "dashed", inherit.aes = FALSE) +
#   scale_color_manual(
#     values = c(setNames(scales::hue_pal()(length(unique(to_gplot$wave))), unique(to_gplot$wave)), "average" = "black"),
#     breaks = c(unique(to_gplot$wave), "average")
#   ) +
#   labs(color = "dataset", shape = 'post-strat table', x = '', y = 'B', subtitle =
#          paste0('results/strat_tables_tmp_approx.Rdata, results/strat_tables_tmp.Rdata, pre.Rmd, ', Sys.Date())) + theme(plot.subtitle=element_text(size=8))
# ggsave('plots/B_alt.png')
```


```{r}
# by table x wave
ggplot(to_gplot, aes(x = wave, y = B, fill = table)) + 
  geom_boxplot() + theme_minimal() + 
  labs(subtitle = paste0('pre.Rmd, ', Sys.Date()))
ggsave('plots/B_wave_x_table.png')

# by dataset 
to_plot %>% group_by(party) %>% summarise(mean(stats.b)) %>% 
  mutate(across(where(is.numeric), round, digits = 4))
```


# correlation score between different methods 

```{r}
# temp data with confidence bounds not 100% correct but approximated
load('results/strat_tables_tmp_approx.Rdata')
out_approx <- out
load('results/strat_tables_tmp.Rdata')
load('results/marginals_raking.Rdata')
to_plot <- rbind(out, out_approx)
method_conv <- c("raking" = 'Raking', "PS-IPW" = "PS-IPW", "post-strat" = 'MN', "DRP" = 'DRP', "MRP" = 'MRP', "NN" = 'NN')
to_plot$method <- method_conv[to_plot$method]
to_plot$method <- factor(to_plot$method, levels = c('Raking', 'PS-IPW', 'MN', 'NN', 'MRP', 'DRP'))
```

```{r}
library(tidyr)
tmp <- to_plot %>% group_by(method, wave, table) %>% summarise(B = mean(B)) # %>% pivot_longer(cols = 'method', values_to = 'B')
round(cor(do.call(cbind, split(tmp$B, tmp$method))), 2)
```

# other statistics such as cell numerosity count 

```{r}
min_cell_size <- 20
out <- data.frame()
for (strat_table in post_strat_tables){
  ref_table <- poptable[[strat_table]]
  for (db in names(main)){
    df <- main[[db]]
    sel_col <- colnames(ref_table)[colnames(ref_table) %in% COLS]
    if (all(sel_col %in% colnames(df))){
      to_plot <- df %>% group_by_at(sel_col) %>% summarise(n = n())
      out <- rbind(out, as.data.frame(t(as.data.frame(c(wave = db, table = strat_table, 
                                                        # false = as.vector(table(to_plot$n < min_cell_size)[1]), 
                                                        true = as.vector(table(to_plot$n < min_cell_size)[2]), 
                                                        nrow(ref_table))))))
    }
  }
}
rownames(out) <- NULL
colnames(out) <- c('wave', 'table', # paste0('cell > ', min_cell_size, ' obs.'), 
                   paste0('cell < ', min_cell_size, ' obs.'), 
                   'n cells')
# colnames(out) <- c('wave', 'table', # paste0('cell = ', min_cell_size - 1, ' obs.'), 
#                    paste0('cell = ', min_cell_size - 1, ' obs.'), 'n cells')
# out$`cell < 20 obs.`[is.na(out$`cell < 20 obs.`)] <- 0
```

```{r}
# create empty dataframe each time 
out_stat <- data.frame()
post_strat_tables <- c('approx', 'edu', 'emp') #, 'pastvote')
for (strat_table in post_strat_tables){
  print(strat_table)
  ref_table <- poptable[[strat_table]]
  for (db in names(main)){
    print(db)
    df <- main[[db]]
    sel_col <- colnames(ref_table)[colnames(ref_table) %in% COLS]
    if (all(sel_col %in% colnames(df))){
      out_method <- NULL
      for (party in coalitions){
        mod_tmp <- glm(paste('nowvote == party ~', paste(sel_col, collapse = ' + ')), data = df, family = binomial(link = 'logit'))
        # proxy strenght 
        out_method[[party]] <- cor(df$nowvote == party, predict(mod_tmp))
      }
      stage2 <- as.data.frame(t(as.data.frame(out_method)))
      colnames(stage2) <- c('proxy str.')
      stage2$levels <- nrow(ref_table)
      sample_dist <- df %>% group_by_at(sel_col) %>% summarise('value' = n() / nrow(df))
      merged <- merge(ref_table, sample_dist, on = sel_col, all = TRUE)
      merged$value[is.na(merged$value)] <- 0
      stage2$mean_diff <- mean(abs(merged$count - merged$value))
      stage2$max_diff <- max(abs(merged$count - merged$value))
    }
    out_stat <- rbind(out_stat, as.data.frame(c(wave = db,
                                  table = strat_table, stage2),
                                stringsAsFactors = FALSE))
    }
}
```

```{r}
stats_out_df %>% group_by(method) %>% summarise(b_str = cor(stats.b, levels), 
                                                B_str = cor(B, levels),
                                                wB_str = cor(wB, levels))
stats_out_df %>% group_by(method) %>% summarise(b_str = cor(stats.b, max_diff), 
                                                B_str = cor(B, max_diff),
                                                wB_str = cor(wB, max_diff))
```


```{r}
# Initialize an empty list to store the results temporarily
results_list <- list()
# Iterate through the combinations of stats and 'b' values
for (stat in c('levels', 'mean_diff', 'max_diff', 'proxy.str.')) {
  for (b in c('stats.b', 'B', 'wB')) {
    # Compute the correlation and store the result
    results_list <- rbind(results_list, data.frame(stat = stat, b = b, 
                                     correlation = cor(stats_out_df[[stat]], stats_out_df[[b]])))
  }
}
# Convert the results list into a dataframe
stage3 <- as.data.frame(results_list)
ggplot(stage3, aes(x = stat, y = b, fill = correlation)) + geom_tile()
```


```{r}
stats_out_df %>% group_by(method) %>% summarise('rho with b stat' = cor(stats.b, proxy.str.), 
                                                'rho with wB' = cor(stats.b, wB),
                                                'rho with B' = cor(stats.b, B)) %>% 
  mutate(across(where(is.numeric), round, digits = 4))
```

```{r}
stage3
```

# time measurement 

```{r}
time_start <- Sys.time()
# generate approximate reference sample
ref_table <- poptable$approx
to_adj <- main$wave4
MN_out <- multinomial(to_adj,
                           ref_table,
                           B = 10
)
time_end <- Sys.time()
(((time_end - time_start) / 10) * 1000) / 60
```

```{r}
time_start <- Sys.time()
# generate approximate reference sample
reference_survey <- generate_reference(census = ref_table)
# logit IPW
ps_ipw(to_adj, reference_survey, trim = 5, B = 10)
time_end <- Sys.time()
(((time_end - time_start) / 10) * 1000) / 60
```


```{r}
time_start <- Sys.time()
# generate approximate reference sample
ref_table <- poptable$approx
to_adj <- main$wave4
reference_survey <- generate_reference(census = ref_table)
# run IPW with a gradient-boosted selection model
xgboost_out <- nonprob_ipw_xgb(
  # non-probability survey
  data = to_adj,
  # generated reference sample
  reference_survey = reference_survey,
  # bootstrap SEs
  var_method = "bootstrap",
  # bootstrap repetitions
  num_boot = 1000
)
time_end <- Sys.time()
time_end - time_start
# around 9 minutes for 1000 iterations in wave 4 
```

```{r}
# NN
time_start <- Sys.time()
ref_table <- poptable$approx
to_adj <- main$wave4
nn_out <- nn_poststrat(
  data = to_adj,
  census = ref_table,
  B = 10, # 1000
  n_epochs = 200,
  n_neurons = 8,
  n_hidden = 2,
  batch_size = 32,
  learning_rate = 1e-3,
  early_stopping = TRUE
)
time_end <- Sys.time()
time_end - time_start #  / 10) * 1000) / 60
(((time_end - time_start) / 10) * 100) / 60
```


```{r}
beepr::beep(3)
```
# ------------------ other code ------------------------

```{r}
coalitions <- c('CSX', 'CDX', 'Az', 'M5S', 'altri')
COLS <-  c('gen', 'age', 'emp', 'edu', 'zone', 'pastvote')
census_list <- c('edu', 'emp', 'pastvote', 'approx')
out <- data.frame(NULL)
for (df_name in names(main)){
  for (census_name in census_list){
    print(paste('processing', df_name, census_name))
    # select census 
    ref_table <- poptable[[census_name]]
    # select data 
    to_adj <- main[[df_name]]
    
    # check or pastvote 
    census_cols <- colnames(ref_table)[colnames(ref_table) %in% COLS]
    data_cols <- colnames(to_adj)[colnames(to_adj) %in% COLS]
    if(all(census_cols %in% data_cols)){
    
      # unweighted
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'unadjusted',
                        uniweighted(to_adj)
      )
      out <- rbind(out, to_append)
      
      # post-stratification or calibration (the same here)
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'post-stratification',
                        poststrat(to_adj, ref_table) 
      )
      out <- rbind(out, to_append)
      
      # multinomial linear model-assisted post-stratification
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'MN post-stratification',
                        multinomial(to_adj, ref_table, 
                                    B = 5 # 1000
                                    )
      )
      out <- rbind(out, to_append)
      
      # generate approximate reference sample
      reference_survey <- generate_reference(census = ref_table)
      # run IPW
      ipw_out <- nonprob_ipw(
        # non-probability survey
        data = to_adj,
        # generated reference sample
        reference_survey = reference_survey, 
        # bootstrap SEs
        var_method = "bootstrap",
        # bootstrap repetitions
        num_boot = 5 # 1000
        # var_method = "analytic" # <- returns the same value of se for all parties
      )
      # logit IPW
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'Logit-IPW',
                        ipw_out
      )
      out <- rbind(out, to_append)
      
      # generate approximate reference sample
      reference_survey <- generate_reference(census = poptable$edu)
      # run IPW with a gradient-boosted selection model
      xgboost_out <- nonprob_ipw_xgb(
        # non-probability survey
        data = to_adj,
        # generated reference sample
        reference_survey = reference_survey,
        # bootstrap SEs
        var_method = "bootstrap",
        # bootstrap repetitions
        num_boot = 2 # 1000
      )
      # XGBoost IPW
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'XGboost-IPW',
                        xgboost_out
      )
      out <- rbind(out, to_append)
      
      # DRP
      to_append <- data.frame(sample = df_name,
                        census = census_name,
                        method = 'DRP',
                        nonprob_drp(to_adj, ref_table) 
      )
      out <- rbind(out, to_append)
      
      # # force rstan to avoid parallel chains
      # options(mc.cores = 1)
      # mrp_out <- mrp(
      #   data = to_adj,
      #   census = ref_table,
      #   draws = 50,
      #   chains = 2,
      #   iter = 200,
      #   cores = 1
      # )
      # # MPR
      # to_append <- data.frame(sample = df_name,
      #                   census = census_name,
      #                   method = 'MRP',
      #                   mrp_out
      # )
      # out <- rbind(out, to_append)
      
      # NN
      # nn_out <- nn_poststrat(
      #   data = to_adj,
      #   census = ref_table,
      #   B = 5,
      #   n_epochs = 50,
      #   n_neurons = 8,
      #   n_hidden = 2,
      #   batch_size = 32,
      #   learning_rate = 1e-3,
      #   early_stopping = TRUE
      # )
      # to_append <- data.frame(sample = df_name,
      #                   census = census_name,
      #                   method = 'NN',
      #                   nn_out
      # )
      # out <- rbind(out, to_append)
      
    }
  }
}
```